---
title: "SI 544 FA 2020 Problem Set 1"
author: "Matthew Garvin"
uniqname: "mgarvin"
output: html_document
---

# SI 544 FA 2020: Problem Set #1

You were introduced to the COVID-19 Data Hub (Guidotti & Ardia, 2020) in this
week's discussion.  We are going to delve a bit deeper into that data set for
this assignment.  We will focus on cleaning and manipulating the data so it's
suitable for generating plots to help us understand the spread of COVID-19.

Google Mobility Reports (GMR) use pooled data to report where people are.
A example of a mobility report can be accessed via
https://www.gstatic.com/covid19/mobility/2020-09-11_US_Mobility_Report_en.pdf.
The GMR provides an indication of the percent change from baseline for the
following types of locations: retail and recreation, grocery and pharmacy,
parks, transit stations, workplaces, and residential.  The GMR data can be
joined to the COVID-19 Data Hub data set by using the `gmr=` option when the
COVID-19 data set is loaded (see first R block below).

In addition to the GMR, the authors of the COVID-19 Data Hub have made it easy
to link the COVID-19 data set with data from the World Bank.  Of particular
interest are two indicators: Total GDP and number of hospital beds per 1000
people.  The first R block below loads country-level data including the two
World Bank indicators mentioned and the second one loads province-level data for
Canada (Canada's provinces are analogous to states in the USA).

```{r load_libraries}
library(tidyverse)
library(COVID19)
```


```{r load_all_countries}
wb <- c("gdp" = "NY.GDP.MKTP.CD", "hosp_beds" = "SH.MED.BEDS.ZS", "diab_prev" = "SH.STA.DIAB.ZS", "water_sanitation" = "IE.PPI.WATR.CD")
all_countries <- covid19(wb = wb)
```

```{r load_canada}
# NOTE: Running this block will take a long time -- be patient!
#       This will also be run when you knit your notebook, 
#       so knitting will also take a few minutes
gmr <- "https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv"
canada <- covid19(country='CAN',gmr = gmr, level=2)
```

To complete this assignment, answer the following questions.
For questions 1-5 use the "all countries" data set (the dataframe named `all_countries`).
Your answers will change depending on when you do your work because the dataset is updated
periodically; keep that in mind when sharing your results with your peers.

1. How many different countries are represented in the country-level data set?
```{r}
# Type your code here
diff_countries <- all_countries %>%
   group_by(administrative_area_level_1) %>%
   count() %>%
   rename(Country = administrative_area_level_1,
          Frequency = n)
```
Answer: 196 different countries


2. What are the earliest and latest dates in the country-level data set?
```{r}
# Type your code here
all_countries %>%
   arrange(date)

all_countries %>%
   arrange(desc(date))
```
Answer: Earliest: 2020-01-01, Latest: 2020-09-22


3. Use the country-level data set to create a histogram that shows the
   distribution of values for the number of hospital beds per 1000 people.  Draw
   a red vertical line that shows the median value on the histogram 
   (hint: use geom_vline(...) as an additional layer in your ggplot).
   (hint: watch out for missing values. You may find `na.rm=TRUE` helpful)
```{r}
# Type your code here

ggplot(data = all_countries, mapping = aes(x = hosp_beds)) +
  geom_histogram(color="white") +
   geom_vline(xintercept = median(all_countries$hosp_beds, na.rm = TRUE), color="red")
median(all_countries$hosp_beds, na.rm = TRUE)
```
Answer: Median = 2.4 hospital beds per 1000 people

4. Use the country-level data set to create a scatterplot of the **total** number of
   recoveries (on the y-axis) vs. the **maximum** number of hospital beds per 1000 people
   (on the x-axis) in that country.
   What does the plot tell you?
```{r}
# Type your code here

dat_filter <- all_countries %>%
   filter(date == "2020-09-20") %>%
   select(id, date, recovered, administrative_area_level_1, hosp_beds) %>%
   summarise(max_bed=replace_na(max(hosp_beds), 0), total_rec=sum(recovered, na.rm = TRUE))
dat_filter


ggplot(data = dat_filter, mapping = aes(max_bed, total_rec)) +
   geom_point(alpha = 0.2)
```
Answer: It seems to indicate that beds matter. Although I'm interested in these outliers in terms of total recovered on the y axis. India has a much larger population, and less than 1 bed per 1000 people, whereas Brazil has around 2 and the U.S. has around 3.


5. Select another World Bank indicator from
   https://data.worldbank.org/indicator/ and create a scatterplot between that
   indicator on the x-axis and the number of confirmed cases (the "confirmed"
   column) on the y-axis. You can download a list of all indicators from
   http://databank.worldbank.org/data/download/site-content/WDI_CETS.xls.
Explain why you chose the indicator you did and provide a brief description of it.

You will need to add the indicator to the list of indicators in the R block that
you use to read in the data.  For example, if you chose "Literacy rate, adult
total (% of people ages 15 and above)?", you would notice that the URL for the
resulting page is
https://data.worldbank.org/indicator/SE.ADT.LITR.ZS?view=chart.  In that case,
the indicator name would be "SE.ADT.LITR.ZS" and your code to read in the data
would be:
```
wb <- c("gdp" = "NY.GDP.MKTP.CD", "hosp_beds" = "SH.MED.BEDS.ZS", "literacy_rate" = "SE.ADT.LITR.ZS")
all_countries <- covid19(wb = wb)
```
Note: you may not use SE.ADT.LITR.ZS as your chosen indicator; choose something else!

----
Answer: I selected diabetes prevalence as a new indicator for the x-axis. I chose this indicator for two reasons: 1) it's a known comordibity of COVID; 2) it is a non-communicable disease associated with increased rates of sedentary lifestyle common in industrialized and developing nations. With this plot I am looking to see if nations with higher rates of diabetes is tied to reduced recovery rates.

This indicator shows diabetes prevalence as a percentage of population between the ages of 20-79

```{r}
# Type your code here

new_dat_filter <- all_countries %>%
   filter(date == "2020-09-21") %>%
   select(id, recovered, diab_prev)
   #summarise(total_rec=sum(recovered, na.rm = TRUE))
new_dat_filter


axs <- ggplot(data = new_dat_filter, mapping = aes(diab_prev, recovered)) +
   geom_point(alpha = 0.2) +
   labs(x = "% of population with diabetes", y = "Total Recovered")
axs
#axs + scale_x_continuous(breaks = c(2000, 4000), labels = c("2k", "4k"))
```

For questions 6-10 use the province-level data set for Canada (the dataframe named `canada`).
Your answers will change depending on when you do your work because the dataset is updated
periodically; keep that in mind when sharing your results with your peers.


6. How many tests have been conducted in Canada in total?
```{r}
# Type your code here
can_dat_filter <- canada %>%
   filter(date == "2020-09-20") %>%
   select(date, key_google_mobility, id, tests) %>%
   summarise(total_tests=sum(tests, na.rm = TRUE))
can_dat_filter

all_tests <- sum(can_dat_filter$total_tests)
all_tests
   
```
Answer: 6,618,405
tests


7. Which Canadian province has conducted the most tests to date?
```{r}
# Type your code here
can_province <- canada %>%
   arrange(desc(tests))
can_province
```
Answer: Ontario

8. Which Canadian province has the highest recovery rate (where recovery rate is
   calculated as the number of recovered cases divided by the number of
   confirmed cases)? Use the most up to date data as you can find using
   "COVID19" package.
```{r}
# Type your code here
can_recover <- canada %>%
   filter(date == "2020-09-19") %>%
   arrange(desc(recovered))
can_recover
   
```
Answer: Quebec

9. Create a bar plot that shows the number of tests conducted per 1000 people
   for each province in Canada. Use the most up to date data as you can find using
   "COVID19" package. (Hint: for Canada, the
   administrative_area_level_2 variable  contains the province name.)
```{r}
# Type your code here

group_province <- canada %>%
   filter(date == "2020-09-20") %>%
   group_by(administrative_area_level_2) %>%
   select(administrative_area_level_2, population, tests) %>%
   mutate(tests_per_1000 = tests / population * 1000)
group_province

group_province %>%
   ggplot(data = group_province, mapping = aes(x = administrative_area_level_2, y = tests_per_1000)) +
   theme(axis.text.x = element_text(angle=60, hjust=1)) +
   geom_col()
```
   
10. Ontario and British Columbia are two provinces in Canada.  Create linegraphs
    that show how the percent change from baseline looks for each of the six
    types of locations (i.e. retail and recreation, grocery and pharmacy, parks,
    transit stations, workplaces, and residential) from the GMR for Ontario. 
    You can experiment with creating one linegraph per location type or, 
    alternatively, you can create multiple lines on one linegraph.
    Repeat this for British Columbia.  

```{r}
# Type your code here
ontario <- canada %>%
   select(date, administrative_area_level_2, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline, parks_percent_change_from_baseline, transit_stations_percent_change_from_baseline, workplaces_percent_change_from_baseline, residential_percent_change_from_baseline) %>%
   filter(id == "83fc0fa9")
ontario

bc <- canada %>%
   select(date, administrative_area_level_2, retail_and_recreation_percent_change_from_baseline, grocery_and_pharmacy_percent_change_from_baseline, parks_percent_change_from_baseline, transit_stations_percent_change_from_baseline, workplaces_percent_change_from_baseline, residential_percent_change_from_baseline) %>%
   filter(id == "c229681f")
bc

retail_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = retail_and_recreation_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = retail_and_recreation_percent_change_from_baseline), color = "blue")
retail_bc_ontario

grocery_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = grocery_and_pharmacy_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = grocery_and_pharmacy_percent_change_from_baseline), color = "blue")
grocery_bc_ontario

parks_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = parks_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = parks_percent_change_from_baseline), color = "blue")
parks_bc_ontario

transit_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = transit_stations_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = transit_stations_percent_change_from_baseline), color = "blue")
transit_bc_ontario

workplaces_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = workplaces_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = workplaces_percent_change_from_baseline), color = "blue")
workplaces_bc_ontario

residential_bc_ontario <- ggplot() +
   geom_line(data = ontario, mapping = aes(x = date, y = residential_percent_change_from_baseline), color = "red") +
   geom_line(data = bc, mapping = aes(x = date, y = residential_percent_change_from_baseline), color = "blue")
residential_bc_ontario

ggplot(data = ontario, mapping = aes(x = date, y = retail_and_recreation_percent_change_from_baseline)) +
   geom_line(color = "red") +
   labs(x = "Date", y = "Ontario percent change from baseline") +
   geom_line(data = ontario, aes(date, grocery_and_pharmacy_percent_change_from_baseline), color = "blue") +
   geom_line(data = ontario, aes(date, parks_percent_change_from_baseline), color = "green") +
   geom_line(data = ontario, aes(date, transit_stations_percent_change_from_baseline), color = "orange") +
   geom_line(data = ontario, aes(date, workplaces_percent_change_from_baseline), color = "purple") +
   geom_line(data = ontario, aes(date, residential_percent_change_from_baseline), color = "black")
ggplot

ggplot(data = bc, mapping = aes(x = date, y = retail_and_recreation_percent_change_from_baseline)) +
   geom_line(color = "red") +
   labs(x = "Date", y = "British Columbia percent change from baseline") +
   geom_line(data = bc, aes(date, grocery_and_pharmacy_percent_change_from_baseline), color = "blue") +
   geom_line(data = bc, aes(date, parks_percent_change_from_baseline), color = "green") +
   geom_line(data = bc, aes(date, transit_stations_percent_change_from_baseline), color = "orange") +
   geom_line(data = bc, aes(date, workplaces_percent_change_from_baseline), color = "purple") +
   geom_line(data = bc, aes(date, residential_percent_change_from_baseline), color = "black")
ggplot
   

```

    Interpret your results: overall, how are
    people behaving differently with COVID-19?  How do people's behaviours
    differ between Ontario and British Columbia?

Answer: More people were staying home in Ontario than in British Columbia. This is particularly apparent in the residential, workplace, and transit plots. More people in British Columbia were out at the parks at least from April - June, while more people in Ontario headed to the parks in July. Ontario shows a deeper decline in retail and grocery plots, but you can also see very similar upticks and downswings among both provinces in these plots.
