---
title: "CH 3-4 Data Wrangling and Tidy Data"
author: "Matt Garvin"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3 Data Wrangling

## Needed Packages

```{r}
library(dplyr)
library(ggplot2)
library(nycflights13)
```
## Filter rows


```{r}
portland_flights <- flights %>%
  filter(dest == "PDX")
View(portland_flights)
```


```{r}
alaska_flights <- flights %>% 
  filter(carrier == "AS")
```


```{r}
btv_sea_flights_fall <- flights %>% 
  filter(origin == "JFK" & (dest == "BTV" | dest == "SEA") & month >= 10)
View(btv_sea_flights_fall)
```


```{r}
not_BTV_SEA <- flights %>% 
  filter(!(dest == "BTV" | dest == "SEA"))
View(not_BTV_SEA)
```

```{r}
flights %>% filter(!dest == "BTV" | dest == "SEA")
```

```{r}
many_airports <- flights %>% 
  filter(dest %in% c("SEA", "SFO", "PDX", "BTV", "BDL"))
View(many_airports)
```


## Learning Check LC3.1
What's another way of using the "not" operator ! to filter only the rows that are not going to Burlington, VT nor Seattle, WA in the flights data frame? Test this out using the previous code.
```{r}
many_airports <- flights %>% 
  filter(!(dest %in% c("SEA", "SFO", "PDX", "BTV", "BDL")))
View(many_airports)
```


## Summarize variables

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp), std_dev = sd(temp))
summary_temp
```

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_temp
```


## Learning Check LC3.2
Say a doctor is studying the effect of smoking on lung cancer for a large number of patients who have records measured at five-year intervals. She notices that a large number of patients have missing data points because the patient has died, so she chooses to ignore these patients in her analysis. What is wrong with this doctor’s approach?

ANSWER: She's completely ignoring deaths related to lung cancer, which is a pretty big effect of smoking.


## LC3.3
Modify the earlier summarize() function code that creates the summary_temp data frame to also use the n() summary function: summarize(... , count = n()). What does the returned value correspond to?
```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            count = n())
summary_temp
```



## LC3.4
Why doesn’t the following code work? Run the code line-by-line instead of all at once, and then look at the data. In other words, run summary_temp <- weather %>% summarize(mean = mean(temp, na.rm = TRUE)) first.


```{r}
summary_temp <- weather %>%   
  summarize(mean = mean(temp, na.rm = TRUE)) %>% 
  summarize(std_dev = sd(temp, na.rm = TRUE))
summary_temp
```


## 3.4 group_by rows

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_monthly_temp
```



