---
title: "Exercises using game data"
author: "Matthew Garvin" "mgarvin"
date: "10/1/2020"
output: html_document
---

# Part 2: Exploring Problem Set 2 Data
```{r}
library(tidyverse)
game_data <- read_csv('bgg_2018.csv')
```

## Explore the "dataset"
By dataset, we mean `game_data` as a dataframe. When exploring, look at the column titles.

## Does there seem to be a relationship between geek_rating and avg_rating?

```{r}
# check correlation
cor(game_data$avg_rating, game_data$geek_rating)
```
Q: What can you say about the relationship based on the output from `cor`?

```{r}
# let's create a scatter plot
ggplot(game_data, aes(y=avg_rating, x=geek_rating))+
  geom_point()
```

```{r}
# how about trying jitter to deal with overplotting?
ggplot(game_data, aes(y=avg_rating, x=geek_rating))+
  geom_jitter(width = 0.25, height = 0.25)
```

```{r}
# let's add a line
ggplot(game_data, aes(y=avg_rating, x=geek_rating))+
  geom_jitter(width = 0.25, height = 0.5)+
  geom_smooth(method=lm)
```
Q: What can you say about the relationship based on the above graph?


## Get top 50 games and check how the avg_rating varies based on certain factors (age and max_players)
```{r}
# get top 50 games
top_50_games <- game_data %>% filter(rank<=50)
```

```{r}
#how does the rating vary with number of players?
ggplot(top_50_games, aes(factor(max_players), avg_rating))+
  geom_boxplot()
```

```{r}
#how does the rating vary with age?
ggplot(top_50_games, aes(factor(age), avg_rating))+
  geom_boxplot()
```

## Get games with 5, 10, and 15 max players and check how the avg_rating varies
```{r}
# get data for games with 5, 10, and 15 max players
game_data_with_specific_players <- game_data %>% filter(max_players %in% c(5, 10, 15))
```

```{r}
# how does the avg_rating vary with number of max_players for games with 5, 10, and 15 max_players?
ggplot(data = game_data_with_specific_players, aes(x = factor(max_players), y = avg_rating))+
  geom_boxplot()
```

```{r}
# simpler steps to get the data and create a boxplot
game_data_5_players <- game_data %>% filter(max_players==5)

ggplot()+
  geom_boxplot(data = game_data_5_players, aes(max_players, avg_rating))

summary(game_data_5_players$avg_rating)

```

Q: From the output of the above written code, map the summary measures to different parts of the box plot.

```{r}
# get data for max_players = 10 and create a box plot
game_data_10_players <- game_data %>% filter(max_players==10)

ggplot()+
  geom_boxplot(data = game_data_10_players, aes(max_players, avg_rating))

summary(game_data_10_players$avg_rating)
```

```{r}
# get data for max_players = 15 and create a box plot
game_data_15_players <- game_data %>% filter(max_players==15)

ggplot()+
  geom_boxplot(data = game_data_15_players, aes(max_players, avg_rating))

summary(game_data_15_players$avg_rating)
```

Q: What are the differences between ratings for games with 5, 10, and 15 max_players?