---
title: "DS6"
date: "10/5/2020"
name: "Matt Garvin"
uniqname: "mgarvin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear regression with categorical variables
* Take away message: `lm(y~factor(x))` is a handy way of comparing means across subgroups in `x`.

```{r}
library(ggplot2)
library(dplyr)
library(moderndive)
df = txhousing
```

## Exercise: Data exploration
Try to answer the following questions using what you have learned so far about R and Rstudio:
1. What is the dataset about? Housing sales in TX
2. What does `median` stand for? Median sale price
3. How is `median` related to `year`? Median sale price is calculated per month in a specific year.


```{r}
# Type your code here
# Q1: What is the dataset about?
help(txhousing)
# Q2: 
# Then, look in the help doc for "median"
# Q3: to get a general sense, plotting a scatter plot can be a good start
ggplot(data = df, mapping = aes(x = year, y = median)) +
    # geom_point()
    geom_jitter()
# Alternative plot
df %>%
    ggplot(aes(x = factor(year), y = median)) +
    geom_boxplot()
```


## Discussion: how to model median sale price as a function of year?
1. Can you predict the `median` sale price for 2016? If yes, what kind of a model would we need? Linear regression model.
2. How are the "estimated means of the `median` (median sale price)" different between years? Looks like a very slight increase from year to year.

We can model the median sale price using year in two ways - 1) treat year as a continuous variable, and 2) treat year as a categorical variable. The objective is to understand the role of these two models in answering the questions above
## `year` as a continous variable
```{r}
lm_fit_years_conti = lm(median ~ year, data=df)
get_regression_table(lm_fit_years_conti)
get_regression_summaries(lm_fit_years_conti)
```


```{r}
lm_fit_years_as_factor = lm(median ~ factor(year), data=df )
#get_regression_summaries(lm_fit_years_as_factor) # Note, this line will fail to knit. Comment it out when knitting for submission.
```

```{r}
df_year_as_cat = df %>% mutate(year = factor(year))
lm_fit_years_as_factor = lm(median ~ year, data=df_year_as_cat )
get_regression_table(lm_fit_years_as_factor)
get_regression_summaries(lm_fit_years_as_factor)
```
In the output above, which year is the reference year? 2000

## What does `y ~ factor(year)` really tells us?

```{r}
mean_of_median_price_by_year = df %>%
    group_by(year) %>%
    summarise(mean_of_median_price = mean(median, na.rm = TRUE))
 
mean_of_reference_group = mean_of_median_price_by_year$mean_of_median_price[1] 

mean_of_median_price_by_year %>%
    mutate(
        the_diff = mean_of_median_price - mean_of_reference_group
    )
```
## Let's alternate the "reference group" and see how the model estimate changes

### Demo: reordering a factor variable
1. We can change how different levels of a factor variable are "ordered" by specifying the order by hand
2. By default, in the `lm()`-regression, the first "group" in the factor variable is taken as the reference group.
```{r}
df %>% mutate(
    # This is one way we can "change the reference group"
    year = factor(year, levels = c(2005, 
                                   2000:2004, 2006:2015 ))
    ) %>%
    ggplot(aes(x = year, y = median)) + 
    geom_boxplot()
```


### Exercise: use the new `year` variable for regressoin
```{r}
# Type your code here:

# Assign the adjusted dataframe to a variable
df_year_as_cat_2005_as_refgroup = df %>% mutate(year = factor(year, levels = c(2005, 2000:2004, 2006:2015)))

#get_regression_table(lm_fit_years_reordered)

# estimate the linear model with year as a categorical variable
lm_fit_years_reordered = lm(median ~ year, data=df_year_as_cat_2005_as_refgroup )

# use get_regression_table() to check what does the 
get_regression_table(lm_fit_years_reordered)
```

## Exercise: again, what does `y ~ factor(year)` really tells us?

```{r}
# Adjust the code from the chunk under section "What does `y ~ factor(year)` really tells us?"
mean_of_median_price_by_year = df_year_as_cat_2005_as_refgroup %>%
    group_by(year) %>%
    summarise(mean_of_median_price = mean(median, na.rm = TRUE))

mean_of_reference_group = mean_of_median_price_by_year$mean_of_median_price[1] 

mean_of_median_price_by_year %>%
    mutate(the_diff = mean_of_median_price - mean_of_reference_group)

```

Going back to the two questions we asked in the beginning:
1. Can you predict the `median` sale price for 2016? If yes, what kind of a model would we need? With lm(median ~ year), if year is counted as a continuous variable, we can calculate a predictive model for 2016.
2. How are the "estimated means of the `median` (median sale price)" different between years? lm(median ~ factor(year)) can help compare means of median across the subgroups of year.

Try writing the answers to these in your own words.

