---
title: "Worksheet for Discussion 2"
author: "Please put down your name here"
uniquename: "Please fill in your uniquename"
output: html_document
---
# Plotting exercises

## Overview:
Over the lecture last week, we went through the following three Known Graphs:
1. Scatter plot
2. Line plot 
3. Histogram

# With everything going on in the world, let data make us feel empowered. Why don't we do COVID plotting?
* Where to get the data?
    * <https://covid19datahub.io/articles/api/r.html>
* What to replicate?
    * CDC COVID Data Tracker - trends:
      <https://covid.cdc.gov/covid-data-tracker/#trends_dailytrends>

## Step 1: Installing the package
R has a COVID19 package that contains data on the status of the pandemic from
around the world for different countries. 

```{r}
# install the package - run the following command (without the "# " piece) in # Console
# install.packages("COVID19")

# load the package
library("COVID19")
```

# Exercise 1. Line graph: new cases over time

## Get data for the US
```{r, echo = TRUE, message=FALSE}
US_cases <- covid19(country = "US")
```

### Explore what we have
```{r}
library(dplyr)
glimpse(US_cases)
```


## Create new a column called "new_cases" for US from the existing data
```{r}
# This chunk will make sense after Friday's lecture
# We create a new column named "new_cases", to represent the increase in the number of confirmed cases on that day
US_cases <- US_cases %>% 
  mutate(confirmed_yesterday = lag(confirmed, n=1), # create column that is previous days value                            
         new_cases = confirmed - confirmed_yesterday) # calculate difference
```

### Explore what we got after adding a new column
```{r}
glimpse(US_cases)
```

### Check that the new_cases column is correct
Subtract the confirmed case counts for two consecutive days to get the number of new_cases for a later date
```{r}
library(knitr)
kable(
    head( US_cases[c( 'date', 'new_cases','confirmed')],
          n=10
    )
)
```

## Get new_cases for Canada
```{r}
Canada_cases <- covid19(country = "CAN")
# Create 
Canada_cases <- Canada_cases %>% 
  mutate(confirmed_yesterday = lag(confirmed, n=1), # create column that is previous days value                            
         new_cases = confirmed - confirmed_yesterday) # calculate difference
```

## Exercise 1: Why is Chris in Canada
In this exercise, we will plot line graphs to compare new cases in US and Canada.

### Part 1: Line graph for new cases in the US
Draw a line plot with `date` on the x-axis and `new_cases` on the y-axis, using the `US_cases` dataframe.
```{r}
# Type your code here
library(ggplot2)
ggplot(US_cases, aes(x = date, y=new_cases))+
  geom_line() 
```

### Part 2: Line graph for new cases in the Canada
```{r}
# Type your code here
ggplot(Canada_cases, aes(x = date, y=new_cases))+
  geom_line() 
```

### Part 3: wait, but why is Chris in Canada?
Try to overlay the two line graphs with the same x-axis. Use different colors to
differentiate the two lines.

```{r}
# Type your code here
# Hint: first, plot one line graph as usual; 
#       then, "add" the second line graph to the one you created in the first step 
#       (recall adding layers from the lecture).

# Solution:
# Plotting US vs Canada
ggplot(Canada_cases,  aes(x = date, y=new_cases))+
  geom_line(color = 'red') + 
  geom_line(data = US_cases, aes(x = date, y = new_cases), color='blue')
```


# Exercise 2: How well are the states testing for COVID19? (Snapshot of Saturday, 2020-09-12)

## Setting things up
Get state-level data from the dataset 
```{r}
US_statelevel = covid19(country = "US", level = 2)
# Check and confirm that the `administrative_area_level_2` column contains the State info;
# These are the available occurrances of each state in the dataset.
table(US_statelevel$administrative_area_level_2)
```

Take a snapshot of 2020-09-12, and plot the histogram of number of tests per 100,000 population:
```{r}
# These will become natural after the lecture on Friday :)
df_2020_09_12 = US_statelevel %>%
  filter(date == "2020-09-12")

# And, plot using "pipes"
df_2020_09_12 %>% 
  mutate(tests_per_100000 = tests / population * 100000) %>%
  ggplot(mapping = aes(x = tests_per_100000)) +
  geom_histogram()
```
From what you plotted above, what is the approx. testing rate for the majority of the states?

## Exercise 2.1: replicate the histogram above
Let's do it step-by-step here: 

### Step 1: construct a dataframe with "tests_per_100000" column
```{r}
# Let's do it step by step
df_Sunday =  df_2020_09_12 
# Pause: check how many columns does df_Sunday have?
df_Sunday = df_Sunday %>% 
  mutate(tests_per_100000 = tests / population * 100000)
# Pause: check again how many columns does df_Sunday have?
```
This is a preview of the `mutate` function: it keeps all columns in the
dataframe and add the ones that are newly created per its arguments.

### Step 2: reproduce the histogram for the distribution of `tests_per_100000`
Note, please use `color=` argument and `binwidth=` argument to make your
histogram more readable.
```{r}
# Type your code here
ggplot(df_Sunday, aes( x = tests_per_100000)) + 
  geom_histogram(color="black", binwidth = 5000)
```
What do you observe? In particular, which bin is the tallest? What does this
imply?


## Exercise 3: construct a bar-plot of tests_per_100000 over all the states
```{r}
# Prep: we need to recode the state variable, as below
df_Sunday$state = df_Sunday$administrative_area_level_2
```

Now, use the newly created `state` variable to plot the bar plot.

```{r}
# Type your code here
ggplot(df_Sunday, aes( x = state, y = tests_per_100000))+
  geom_col() +
  labs(caption="A bar chart of testing rate for different states")
```

Can you read the text labels on the x-axis, for all the states? If not, try to
"add" the following code towards the end of your chunk:
```
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
```

```{r}
#We can improve the look and feel of the labels using the theme option
ggplot(df_Sunday, aes( x = state, y = tests_per_100000))+
  geom_col() +
  labs(caption="A bar chart of testing rate for different states") +
  theme(axis.text.x = element_text(angle=60, hjust=1))
```

From the histogram, we saw that a majority of states are testing around 20,000
out of 100,000 (or 2 out of 10). Can we confirm this in the bar plot?
```{r}
# Here, we draw a line at 20000 and check if what the histogram told us makes sense
# To begin with, we draw the same bar plot using geom_col:
ggplot(df_Sunday, aes( x = state, y = tests_per_100000))+
  geom_col() +
  labs(caption="A bar chart of testing rate for different states") +
  theme(axis.text.x = element_text(angle=60, hjust=1)) +
# Lastly, we add the horizontal line
  geom_abline(slope=0, intercept = 20000, color='red') #draw a horizontal line (slope=0)
```
