---
title: "Week1_Discussion_WithSolution"
author: "Matt Garvin"
date: "9/10/2020"
output: html_document
---

# Introduction
This in-class notebook is designed to get you used to working with R and
Rstudio. People learn through play, so feel free to play around. Experiment.
Try things. If you end up breaking R, you can just restart the kernel.

If you find this easy, be willing to help classmates during breakout sessions.
If you find it difficult, don't be shy about asking your neighbors for help.

## Learning objectives (pending)
1. Get R and Rstudio installed, and install packages.
2. Get used to getting `help` with `help()` function.
3. Create and work with variables, vectors.

# Demo

## Prepare your R environment
1. Create a new R Markdown file through the menu: `File/New File/R Markdown`. 
  This shall set your R+Rstudio installation ready for manipulating the `*.Rmd` 
  notebook files.
2. Try to install and load packages, as introduced in the 
   [Chapter 1](https://moderndive.com/1-getting-started.html) of the textbook.
   Execute the following command in the Console (Ctrl+2) - same for MAC and WINDOWS
  ```
  install.packages("dplyr")
  install.packages("knitr")
  ```
3. "Knit" the Rmd document into HTML.
    * Keyboard Shortcut: `Ctrl+Shift+K`,
    * Or, through the Menu: `File - Knit Document`.
    * Or, through the Knit icon at the top of this pane.

### How to not knit the document
1. Delete a backtick sign from "chunks" (backtick is the key to the left of number 1, on most keyboards)
2. Typos in the header of the chunk: `FALSE` as `False` will fail us.
3. For macOS, missing the [XQuartz](https://www.xquartz.org/) program (which is a X-server).

# R basics

## Special notes
* In R, `TRUE` and `FALSE` denote the binary values. This is different from the
  `True` and `False` spellings in Python.
* Naming conventions: 
    1. Object names cannot begin with a number.
    2. Object names cannot contain symbols used for mathematics or to denote
       other operations native to R. These symbols include 
       `$`, `@`, `!`, `^`, `+`, `-`, `/`, and `*`.

## Assigning values to variables
You can assign single values to elements.
    ```{r}
    # You can assign a single value
    my_variable <- 7.3
    # You can even assign an irrational number
    v <- pi
    # You can create vectors using the c() function
    vec <- c(5, 3, 0.2, -20)
    ```

What happens in the Environment pane as we execute the command below?


### Exercise 1: Create the following variables.

Ensure that you can see them in the
**Environment** tab, which should be on the top right of the RStudio
window.

  Variable Name                   Value
  ------------------------------- ------------------
  x                               2
  y                               25, 0, -4, 0.09
  vec1                            1, 2, 3, 4, 5, 6
  super.long.variable.name        0
  another_long_variable_name      1

    ```{r}
    x <- 2
    y <- c(25, 0, -4, 0.09)
    vec1 <- c(1, 2, 3, 4, 5, 6)
    super.long.variable.name <- 0
    another_long_variable_name <- 1
    ```

## Basic Operations

For each of the following expressions **FIRST write down what you
think R will give you (make a guess).** Then see what R actually does
give you. Make sure you understand why it gave you the answer it did.

(A) y-x 
(B) y+5
(C) vec1*x 
(D) vec1/super.long.variable.name
(E) c(y, vec1)
(F) vec1 + y 
(G) vec1*vec1
(H) sqrt(y) 
(I) sum(y)



    ```{r}
    y - x
    y + 5
    vec1 * x
    vec1/super.long.variable.name
    c(y, vec1)
    vec1 + y
    vec1*vec1
    sqrt(y)
    sum(y)
    ```

## Vectors
What do you get from expression `4:14`? How about `c(4:14)`? Are they the same?
    ```{r}
    4:14
    c(4:14)
    4:14==c(4:14) # Why this?
    # Then, assign into variables and check if they are "identical"
    v1 <- 4:14
    v2 <- c(4:14)
    identical(v1, v2) # How is this different from 4:14==c(4:14)?
    ```

### Exercise:
What expression will generate the following vector?
    ```
    -3, -2, -1, 0, 1, 2, 3, 4
    ```

Please type your code here. [Try to create a new "chunk" here yourself :)]
```{r}
c(-3, -2, -1, 0, 1, 2, 3, 4)
```


### Compute the average
`mean()` function shall take the average. What's the average of numbers in
`vec1`?

```{r}
mean(vec1)
```

## Data exploration
R comes loaded with some data sets. For example, `cars`. What is the cars data
about? Type `help(cars)` to read more about it.
```{r}
help(cars)
```

Let's assign it to a new variable called `carsdata`:

```{r}
carsdata <- cars
```




### Exercise: explore the `carsdata` dataframe
Use  `View()`, `glimpse()` and `kable()` to explore the `cars` dataframe.
Note, the `glimpse()` function is a available in the package `dplyr`. And, the
`kable()` function is available in package `knitr`.
    * Try writing `dplyr` in your console next to the prompt and see what happens.
Reference: [Section 1.4.3 in ModernDive](https://moderndive.com/1-getting-started.html#exploredataframes)

    ```{r echo=TRUE}
    # First, we need to load the packages we need
    library(dplyr)
    library(knitr)
    ```

    ```
    # Please copy this to the console and execute it (press enter)
    View(carsdata)
    ```

    ```{r}
    glimpse(carsdata) # glimpse is from dplyr
    ```

    ```{r}
    kable(carsdata) #kable is from knitr
    ```

    What does `carsdata$speed` get you? Run the code below to find out 
    ```{r}
    carsdata$speed
    ```

### Exercise: convert to Canadian standards
Create a new `speed_c` column based on the `speed` column, where the unit is km/h. (1 mph = 1.609344 km/h)

    ```{r}
    # Your code goes here
carsdata$speed_c = carsdata$speed * 1.609344
    ```
Create a new `dist_c` column based on the `dist` column, where the unit is meters. (1 ft = 1/3.281 m )

    ```{r}
    # Your code goes here
carsdata$dist_c = carsdata$dist * 1/3.281
    ```
Review the dataframe using `View(carsdata)`: we created two new columns.

# Submission
Please "knit" this R Markdown file and submit the compiled HTML file to
Canvas. You will find an entry for Discussion Section 1 the Assignments section on Canvas.
