---
title: "Counting a simulated Normal distribution + Z-score from sample means"
author: ""
date: ""
---

# Part 1: simulating normal distribution and calculating the quantiles



## Simulation: random sample from Normal(9, 30)
In the following code snippet, we generate a random sample from a Normal distribution with `mean = 9` and `standard_deviation=30`.


```{r}
library(ggplot2)

the_mean <- 9
the_standard_deviation <- 30

set.seed(1) #Q: why we set a seed? For what?
sim_count = 1000000 # generate a random sample of size a million

df <- data.frame(v1 = rnorm(sim_count,
                            mean = the_mean, 
                            sd = the_standard_deviation 
                            ) 
                 )
summary(df)
```

### Interlude: why we set the seed? For reproducibility.
Source: <http://rfunction.com/archives/62>
```{r}
set.seed(5)
print(rnorm(5)) # A batch of random numbers
print(rnorm(5)) # Another batch of random numbers
```


```{r}
set.seed(5)
print(rnorm(5)) # The same batch of random numbers
```


### Calculate the z-score for v1
```{r}
df$z_score = (df$v1 - mean(df$v1)) / sd(df$v1)
summary(df)
```



### Inspect the histogram for `df$v1` and `df$z_score`

```{r}
ggplot(df) + 
    geom_histogram(aes(x = v1, ),
                   colour = "black",
                   bins = 100,
                   fill = "blue") + 
    geom_histogram(aes(x = z_score, ),
                   colour = "black",
                   bins = 100,
                   fill = "white") 
```
### And, a closer look at the histogram for the z_score:
```{r}
ggplot(df) + 
    geom_histogram(aes(x = z_score, ),
                   colour = "black",
                   bins = 100,
                   fill = "white") 
```

## Density plot: the area under the graph is 1.

```{r}
ggplot(df, aes(x = z_score)) + 
geom_histogram(aes(y =..density..),
               colour = "black",
               bins = 100,
               fill = "white") +
    stat_function(fun = dnorm, args = list(mean = mean(df$z_score), sd = sd(df$z_score)), color="red")
```

### Draw auxiliary lines at 0, 1, 2

```{r}
ggplot(df, aes(x = z_score)) + 
geom_histogram(aes(y =..density..),
               colour = "black",
               bins = 100,
               fill = "white") +
    stat_function(fun = dnorm, args = list(mean = mean(df$z_score), sd = sd(df$z_score)), color="red") +
geom_vline(xintercept = 0, color='black') +
    geom_vline(xintercept = sd(df$z_score), color='blue') +
    geom_vline(xintercept = - sd(df$z_score), color='blue') +
    geom_vline(xintercept = sd(df$z_score) * 2, color='red') +
    geom_vline(xintercept = - sd(df$z_score) * 2, color='red')
    ```
The following statements are equivalent given that there are 1,000,000
1. By count: There are roughly 950,000 random samples whose z_score value is
   between the two red lines;
2. By "probability": out of a total of 1,000,000 values in vector df$z_score,
   95% of them lie between the two red lines
3. By area: The area under the curve that is between the two red lines is 0.95

## Counting vs probability
For values stored in `df$z_score`, what is the fraction of numbers that are
smaller than its mean?
```{r}
# calculate the mean:
the_mean <- mean(df$z_score)
the_mean # This is very close to 0.
```

```{r}
library(dplyr)
# Count the number of rows that has z_score smaller than the mean
count_smaller_than_mean = df %>% filter(z_score <= the_mean) %>%
    summarise(count = n()) 
count_smaller_than_mean$count
```

Then, what is the probability that a randomly chosen number from `df$z_score` is
smaller than its mean, 0?
```{r}
count_smaller_than_mean$count / length(df$z_score)
```

### Exercise
For values stored in `df$z_score`, what is the fraction of numbers that are
smaller than two times its standard deviation?
```{r}
# Step 1: get the value "two times its standard deviation"
value_2sd = 2 * sd(df$z_score)
```

```{r}
# Step 2: Count
count_smaller_than_2sd = df %>% filter(z_score <= value_2sd) %>%
    summarise(count = n()) 
count_smaller_than_2sd
```

```{r}
# Step 3: calculate the fraction/probability

count_smaller_than_2sd / length(df$z_score)
```




## Quantiles: the "tick" at which x% are smaller than it
```{r}
# What does "3rd Qu." mean?
summary(df$z_score)
```

### Use df$z_score, calculate the probability that a random chosen number for this column is smaller than its 3rd Quantile
```{r}
count_between_1sd = df %>% filter(z_score <= 0.675059) %>%
    summarise(count = n()) 
count_between_1sd$count / length(df$z_score)
```
That is, there are 75% random samples that are smaller than the 3rd quantile, 0.675059.

## Confidence intervals: 1 standard deviations away from the mean
```{r}
library(dplyr)
count_between_1sd = df %>% filter(
    z_score <= 1 * sd(z_score) & z_score >= -1 * sd(z_score)) %>%
    summarise(count = n())
count_between_1sd$count / sim_count
```

    
## Review: Box plot and histogram
Open [this link](https://commons.wikimedia.org/wiki/File:Boxplot_vs_PDF.svg), or inspect the following plot:

![boxplot and histogram, source: Wikimedia](https://upload.wikimedia.org/wikipedia/commons/1/1a/Boxplot_vs_PDF.svg)



