---
title: "si544-day7-multiple-regression"
author: "Your Name (uniqname)"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As usual, load all the libraries that we'll need:

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
```

Simplify our dataset:

```{r}
evals_day7 <- evals %>% select(ID, score, age, gender)
```

Create a visualization of a linear regression:
```{r}
evals_day7 %>% ggplot(mapping=aes(x=age,y=score)) + geom_point() + geom_smooth(method="lm",se = FALSE)
```

Add the confidence interval:

```{r}
evals_day7 %>% ggplot(mapping=aes(x=age,y=score)) + geom_point() + geom_smooth(method="lm",se = TRUE)
```

Create a linear model of score vs. age using a pipeline:

```{r}
evals_model <- evals_day7 %>% lm(score ~ age, data = .) 
```


```{r}
get_regression_table(evals_model)
```
Now let's separate male and female:

```{r}
ggplot(evals_day7, aes(x = age, y = score, color = gender)) +
  geom_point() +
  labs(x = "Age", y = "Teaching Score", color = "Gender") +
  geom_smooth(method="lm",se=FALSE)
```

And let's create a linear model of score vs. age that takes gender into account:

```{r}
score_model_interaction <- lm(score ~ age * gender, data = evals_day7)
# Get regression table:
get_regression_table(score_model_interaction)
```

Let's look at the residuals:

```{r}
get_regression_points(score_model_interaction)
```

Let's take a look at a model that assumes there is no interaction:

```{r}

gg_parallel_slopes(y = "score", num_x = "age", cat_x = "gender", 
                   data = evals_day7)
```

Let's create a regression model that does not have an interaction term (note the formula
uses '+' instead of '*'):

```{r}
score_model_parallel_slopes <- lm(score ~ age + gender, data = evals_day7)
get_regression_table(score_model_parallel_slopes)
```



```{r}
library(ISLR)
```

```{r}
credit = Credit
```
```{r}
library(skimr)
```

```{r}
skim(credit)
```



Credit Limit Modeling

Credit balance vs. income
```{r}
income_model <- lm(Balance ~ Income, data = credit)
get_regression_summaries(income_model)
get_regression_table(income_model)

```

Credit balance vs. credit rating
```{r}
rating_model <- lm(Balance ~ Rating, data = credit)
get_regression_summaries(rating_model)
get_regression_table(rating_model)
```
Credit balance vs. income and limit

```{r}
income_balance_interaction_model <- lm(Balance ~ Income * Limit, data = credit)
get_regression_summaries(income_balance_interaction_model)
get_regression_table(income_balance_interaction_model)
```

Credit balance vs. income and credit rating and limit

```{r}
balance_rating_interaction_model <- lm(Balance ~ Income * Rating * Limit, data = credit)
get_regression_summaries(balance_rating_interaction_model)
get_regression_table(balance_rating_interaction_model)
```











