---
title: 'SI 544 Day 2: Visualization'
author: "Matt Garvin (mgarvin)"
date: "9/1/2020"
output:
  pdf_document: default
  html_document: default
---

## Day 2: Libraries

Before we get started, let's load the libraries that we'll need for today:

```{r loadlibraries}

library(nycflights13)
library(ggplot2)
library(dplyr)
library(gapminder)
library(knitr)
library(readr)

```

## Gapminder

Here's the code to load the Gapminder data.  
Note that we are filtering the data so we only get the data from the year 2007, and then we drop the year column (why?).
We also rename some columns so that they're easier to deal 
with in the code. For example, we use lowercase for the variable names and we also drop the spaces in the variable names.


```{r gapminder}

gapminder_2007 <- gapminder %>% 
  filter(year == 2007) %>% 
  select(-year) %>% 
  rename(
    Country = country,
    Continent = continent,
    `Life Expectancy` = lifeExp,
    `Population` = pop,
    `GDP per Capita` = gdpPercap
  )
```

You might want to View(gapminder_2007) in the Console.

```{r}
ggplot(data = gapminder_2007, 
       mapping = aes(x = `GDP per Capita`, 
                     y = `Life Expectancy`, 
                     size = Population, 
                     color = Continent)) +
  geom_point() +
  labs(title="Gapminder demo", x = "GDP per capita", y = "Life expectancy")
```
```{r}
alaska_flights <- flights %>% 
  filter(carrier == "AS")
```


```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + geom_point()
```
```{r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(binwidth = 5, color = "white") + facet_wrap(~ month)
```
```{r}
ggplot(data = weather, mapping = aes(x = month, y = temp)) + geom_boxplot()
```

```{r}
ggplot(data = alaska_flights, mapping = aes(x = sched_dep_time, y = sched_arr_time)) + geom_point()
```

```{r}
ggplot(data = alaska_flights, mapping = aes(x = distance, y = air_time)) + geom_boxplot()
```

```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + geom_point(alpha = 0.2)
```

```{r}
ggplot(data = alaska_flights, mapping = aes(x = dep_delay, y = arr_delay)) + geom_jitter(width = 30, height = 30)
```
``` {r}
ggplot(data = early_january_weather, mapping = aes(x = time_hour, y = temp)) + geom_line()
```

``` {r}
ggplot(data = early_january_weather, mapping = aes(x = time_hour, y = visib)) + geom_line()
```

``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram()
```

``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(color = "white")
```

``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(color = "white", fill = "steelblue")
```

``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(bins = 40, color = "white")
```

``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(binwidth = 10, color = "white")
```
``` {r}
ggplot(data = weather, mapping = aes(x = temp)) + geom_histogram(binwidth = 5, color = "white") + facet_wrap(~ month, nrow = 4)
```

``` {r}
ggplot(data = weather, mapping = aes(x = month, y = temp)) + geom_boxplot()
```

``` {r}
ggplot(data = weather, mapping = aes(x = factor(month), y = temp)) + geom_boxplot()
```

``` {r}
fruits <- tibble(fruit = c("apple", "apple", "orange", "apple", "orange"))
fruits_counted <- tibble(fruit = c("apple", "orange"), number = c(3, 2))
ggplot(data = fruits, mapping = aes(x = fruit)) + geom_bar()
```

``` {r}
ggplot(data = fruits_counted, mapping = aes(x = fruit, y = number)) +
  geom_col()
```
``` {r}
ggplot(data = flights, mapping = aes(x = carrier)) +
  geom_bar()
```

``` {r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar()
```


``` {r}
ggplot(data = flights, mapping = aes(x = carrier, color = origin)) +
  geom_bar()
```

``` {r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = "dodge")
```

``` {r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar(position = position_dodge(preserve = "single"))
```

``` {r}
ggplot(data = flights, mapping = aes(x = carrier, fill = origin)) +
  geom_bar() + facet_wrap(~ origin, ncol = 1)
```
``` {r}
ggplot(data = flights, mapping = aes(x = carrier)) + geom_bar()
```

``` {r}
ggplot(flights, aes(x = carrier)) + geom_bar()
```
``` {r}
flights_counted <- flights %>% group_by(carrier) %>% count()
```