---
title: "si544-day5-basic-regression"
author: "Chris Teplovs"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
```

## Load & Look


```{r}
evals_ch6 <- evals %>%
  select(ID, score, bty_avg, age)
```

```glimpse``` can be helpful:

```{r}
glimpse(evals_ch6)
```

Alternatively, you can draw a random sample of some values and look at what you get:

```{r}
evals_ch6 %>% sample_n(5)
```

Now that you understand descriptive statistics and visualizations, let's take a shortcut:


```{r}
evals_ch6 %>%
  select(score, bty_avg) %>%
  skim()

```

```{r}
evals_ch6 %>%
  get_correlation(formula = score ~ bty_avg)
```

Equivalently:

```{r}
evals_ch6 %>%
  summarize(correlation = cor(score, bty_avg))
```

Visualization:

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```


## Simple Linear Regression

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch6)
```


```{r}
# Get regression table:
get_regression_table(score_model)
```

```{r}
get_regression_summaries(score_model)
```

### Best fit

```{r}
regression_points <- get_regression_points(score_model)
regression_points
```

Let's calculate the sum of squared residuals using a pipeline:

```{r}
regression_points %>%
  mutate(squared_residuals = residual^2) %>%
  summarize(sum_of_squared_residuals = sum(squared_residuals))

```


```{r}
ggplot(regression_points, mapping=aes(x=residual)) + geom_histogram()
```



## One categorical predictor variable

Let's return to the Gapminder data:

```{r}
library(gapminder)
gapminder2007 <- gapminder %>%
  filter(year == 2007) %>%
  select(country, lifeExp, continent, gdpPercap)

```


What did the above code block do?

INSERT YOUR ANSWER HERE

```{r}
glimpse(gapminder2007)
```


Use skim() to get a "feel" for lifeExp and continent.  What does the output tell you?

```{r}
# INSERT YOUR R CODE HERE
```

```{r}
ggplot(gapminder2007, aes(x = lifeExp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Life expectancy", y = "Number of countries",
       title = "Histogram of distribution of worldwide life expectancies") +
  facet_wrap(~ continent, nrow = 2)
```

```{r}
ggplot(gapminder2007, aes(x = continent, y = lifeExp)) +
  geom_boxplot() +
  labs(x = "Continent", y = "Life expectancy (years)",
       title = "Life expectancy by continent")
```

```{r}
lifeExp_by_continent <- gapminder2007 %>%
  group_by(continent) %>%
  summarize(median = median(lifeExp), mean = mean(lifeExp))
lifeExp_by_continent
```

