
---
title: "si544-day3-data-wrangling-prerecorded"
author: "Chris Teplovs"
date: "9/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, let's load the libraries we'll need:

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(nycflights13)
library(fivethirtyeight)
```

## Pipes

```{r}
values <- c(1,2,3)
```

Usual approach is to pass the data frame as an argument, that is:

```{r}
sum(values)
```

Using pipes, we can say "take values then apply sum":

```{r}
values %>% sum
```

## Filtering

(You might want to View(flights) in the Console.)

```{r}
alaska_flights <- flights %>%
   filter(carrier == "AS")
```



```{r}
portland_flights <- flights %>%
   filter(dest == "PDX")
```

```{r}
btv_sea_flights_fall <- flights %>%
   filter(origin == "JFK" & (dest == "BTV" | dest == "SEA") & month >= 10)
```

We can use ',' instead of '&':

```{r}
btv_sea_flights_fall <- flights %>%
   filter(origin == "JFK", (dest == "BTV" | dest == "SEA"), month >= 10)
```

Lots of '|' (or) operators:
```{r}
many_airports <- flights %>%
 filter(dest == "BTV" | dest == "SEA" | dest == "PDX" |
        dest == "SFO" | dest == "BDL")
```

The '%in%' operator can replace lots of '|' operators:
```{r}
many_airports <- flights %>%
 filter(dest %in% c("BTV", "SEA", "PDX", "SFO", "BDL"))
```

# Summarization


```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp), std_dev = sd(temp))
summary_temp
```

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_temp
```

# Grouping

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_monthly_temp

```

Note that group_by doesn't change the data.frame, only the information about the data.frame:

```{r}
diamonds
```


```{r}
diamonds %>% 
  group_by(cut)
```

```{r}
diamonds %>% 
  group_by(cut) %>% 
  summarize(avg_price = mean(price))
```


```{r}
by_origin <- flights %>% 
  group_by(origin) %>% 
  summarize(flight_count = n())
by_origin
```

Grouping can also be done for multiple (nested) variables.  For example to group by origin and then by month:

```{r}
by_origin_monthly <- flights %>% 
  group_by(origin, month) %>% 
  summarize(count = n())
by_origin_monthly
```

Note that the following will not work properly:

```{r}

by_origin_monthly_incorrect <- flights %>% 
  group_by(origin) %>% 
  group_by(month) %>% 
  summarize(count = n())
by_origin_monthly_incorrect
```

## Mutating

Creating new variables based on existing ones.

```{r}
weather <- weather %>% 
  mutate(temp_in_C = (temp - 32) / 1.8)
```
```{r}
library(magrittr)
```
```{r}
weather <- weather %>% 
  mutate(temp_in_C = temp %>% subtract(32) %>% divide_by(1.8))
```

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean_temp_in_F = mean(temp, na.rm = TRUE), 
            mean_temp_in_C = mean(temp_in_C, na.rm = TRUE))
summary_monthly_temp
```

Calculate a new variable "gain" that is the the difference of dep_delay and arr_delay:

```{r}
flights <- flights %>% 
  mutate(gain = dep_delay - arr_delay)
```



```{r}
gain_summary <- flights %>% 
  summarize(
    min = min(gain, na.rm = TRUE),
    q1 = quantile(gain, 0.25, na.rm = TRUE),
    median = quantile(gain, 0.5, na.rm = TRUE),
    q3 = quantile(gain, 0.75, na.rm = TRUE),
    max = max(gain, na.rm = TRUE),
    mean = mean(gain, na.rm = TRUE),
    sd = sd(gain, na.rm = TRUE),
    missing = sum(is.na(gain))
  )
gain_summary
```


```{r}
ggplot(data = flights, mapping = aes(x = gain)) +
  geom_histogram(color = "white", bins = 20) +
  labs(caption="Histogram of gain variable")
```



```{r}
flights <- flights %>% 
  mutate(
    gain = dep_delay - arr_delay,
    hours = air_time / 60,
    gain_per_hour = gain / hours
  )
```



# Sorting (arrange)

```{r}
freq_dest <- flights %>% 
  group_by(dest) %>% 
  summarize(num_flights = n())
freq_dest
```

```{r}
freq_dest %>% 
  arrange(num_flights)
```

```{r}
freq_dest %>% 
  arrange(desc(num_flights))
```



## Joining (inner_join)


```{r}
flights_joined <- flights %>%
 inner_join(airlines, by = "carrier")
``` 





```{r}
flights_with_airport_names <- flights %>%
 inner_join(airports, by = c("dest" = "faa"))
``` 


```{r}
named_dests <- flights %>%
  group_by(dest) %>%
  summarize(num_flights = n()) %>%
  arrange(desc(num_flights)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  rename(airport_name = name)
named_dests
```

```{r}
flights_weather_joined <- flights %>%
 inner_join(weather, by = c("year", "month", "day", "hour", "origin"))
```

```{r}
joined_flights <- flights %>%
 inner_join(airlines, by = "carrier")
```


## Selecting

```{r}
flights %>%
 select(carrier, flight)
```

```{r}
flights_no_year <- flights %>%
 select(-year)
```


```{r}
flight_arr_times <- flights %>%
 select(month:day, arr_time:sched_arr_time)
```


```{r}
flights_reorder <- flights %>%
 select(year, month, day, hour, minute, time_hour, everything())
```

```{r}
flights_begin_a <- flights %>%
 select(starts_with("a"))
```

```{r}
flights_delays <- flights %>%
  select(ends_with("delay"))
```


```{r}
flights_time <- flights %>%
 select(contains("time"))
```

## Renaming

```{r}
flights_time_new <- flights %>%
 select(contains("time")) %>%
 rename(departure_time = dep_time,
        arrival_time = arr_time)
```


## Importing

We saw last week that we can load a CSV file:

```{r}
q <- read.csv('544fa2019questionnaire.csv')
```

Note that we can also read directly from a URL, in this case it's a CSV published from a Google Sheets file:

```{r}
hw <- read.csv('https://bit.ly/2kPxxS3')
```

Note that R can also read Excel files directly.  See the textbook for details.

## Tidying

```{r}
library(fivethirtyeight)
```

```{r}
drinks
```

```{r}
drinks_smaller <- drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia")) %>% 
  select(-total_litres_of_pure_alcohol) %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings)
drinks_smaller
```

```{r}
drinks_smaller_tidy <- drinks_smaller %>% 
  gather(key = type, value = servings, c(beer, spirit, wine))
drinks_smaller_tidy
```

```{r}
ggplot(drinks_smaller_tidy, 
       aes(x = country, y = servings, fill = type)) +
  geom_col(position = "dodge")
```

END OF NOTEBOOK

``` {r}
library(corrgram)
library(corrplot)
#creating 'hours' column and making sure that hours are only from
#00:00 hrs to 23:00 hrs which is exactly as the 'weather' data set has.
#this will help us join the 2 data sets properly.
flights$hour <- ifelse(flights$hour == 24, 0, flights$hour)

#JOining the 'flights' and 'weather' datasets based on unique identifiers.
flights_weather <- left_join(flights, weather)

#We create a 'delays' column that is addition of all the delays in
#arrival and departure (Consider only positive delays. Reason stated in
#1st problem solution)
flights_weather$arr_delay <- ifelse(flights_weather$arr_delay >= 0,
                                    flights_weather$arr_delay, 0)
flights_weather$dep_delay <- ifelse(flights_weather$dep_delay >= 0,
                                    flights_weather$dep_delay, 0)
flights_weather$total_delay <- flights_weather$arr_delay + flights_weather$dep_delay

 
#creating a data with only delay and weather columns. Removing origin, date/time
#because we want a correlation between delay and the particular weather condition.
cor_data <- select(flights_weather, total_delay, temp, dewp, humid,
                   wind_dir, wind_speed, wind_gust, precip, pressure, visib)

#WE first plot a correlation Matrix using corrplot to find the variables that
#are correlated. We create a correlation matrix using 'cor' function
corrplot(cor(na.omit(cor_data)), method = "circle", type = "upper",
         tl.srt = 25, tl.col = "Black", tl.cex = 1, title = "Correlation
         between all 'weather' variables & 'delay'", mar =c(0, 0, 4, 0) + 0.1)
```
