---
title: "si544-day3-data-wrangling-inclass"
author: "Matt Garvin"
uniqname: "mgarvin"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, let's load the libraries we'll need (missing libraries?  use install.packages() to install them:

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(nycflights13)
library(fivethirtyeight)
```

## Pipes

```{r}
values <- c(1,2,3)
```

The usual approach is to pass the data frame as an argument, that is:

```{r}
sum(values)
```

Using pipes, we can say "take values then apply sum":

```{r}
values %>% sum
```

## Filtering

(You might want to View(flights) in the Console.)

```{r}
alaska_flights <- flights %>%
   filter(carrier == "AS")
```

```{r}
portland_flights <- flights %>%
   filter(dest == "PDX")
```

```{r}
btv_sea_flights_fall <- flights %>%
   filter(origin == "JFK" & (dest == "BTV" | dest == "SEA") & month >= 10)
```

We can use ',' instead of '&':

```{r}
btv_sea_flights_fall <- flights %>%
   filter(origin == "JFK", (dest == "BTV" | dest == "SEA"), month >= 10)
```


```{r}
not_BTV_SEA <- flights %>%
 filter(!(dest == "BTV" | dest == "SEA"))
```

Oops:


```{r}
flights %>%
 filter(!dest == "BTV" | dest == "SEA")
```

Lots of '|' (or) operators:
```{r}
many_airports <- flights %>%
 filter(dest == "BTV" | dest == "SEA" | dest == "PDX" |
        dest == "SFO" | dest == "BDL")
```

The '%in%' operator can replace lots of '|' operators:
```{r}
many_airports <- flights %>%
 filter(dest %in% c("BTV", "SEA", "PDX", "SFO", "BDL"))
```

## (LC 544.3.1)
Write code to create a dataframe containing only those flights arriving at Detroit (airport code DTW)

```{r}
# INSERT YOUR R CODE HERE
detroit_flights <- flights %>% filter(dest == "DTW")
```



# Summarization


```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp), std_dev = sd(temp))
summary_temp
```

```{r}
summary_temp <- weather %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_temp
```

# Grouping

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean = mean(temp, na.rm = TRUE), 
            std_dev = sd(temp, na.rm = TRUE))
summary_monthly_temp

```

Note that group_by doesn't change the data.frame, only the information about the data.frame:

```{r}
diamonds
```


```{r}
diamonds %>% 
  group_by(cut)
```

```{r}
diamonds %>% 
  group_by(cut) %>% 
  summarize(avg_price = mean(price))
```


```{r}
by_origin <- flights %>% 
  group_by(origin) %>% 
  summarize(count = n())
by_origin
```

Grouping can also be done for multiple (nested) variables.  For example to group by origin and then by month:

```{r}
by_origin_monthly <- flights %>% 
  group_by(origin, month) %>% 
  summarize(count = n())
by_origin_monthly
```

Note that the following will not work properly:

```{r}

by_origin_monthly_incorrect <- flights %>% 
  group_by(origin) %>% 
  group_by(month) %>% 
  summarize(count = n())
by_origin_monthly_incorrect
```

## (LC4.5) (LC == Learning Check)
Recall from last class's notebook when we looked at plots of temperatures by months in NYC (Section ## Learning Check (LC3.23) ) . What does the standard deviation column in the summary_monthly_temp data frame tell us about temperatures in New York City throughout the year?

INSERT YOUR ANSWER HERE; DON'T OVERTHINK THIS QUESTION!
That the deviation between temperatures in a given month is relatively close to the mean. It appears the largest deviation is in January at 10.22.

## Mutating

Creating new variables based on existing ones.

```{r}
weather <- weather %>% 
  mutate(temp_in_C = (temp - 32) / 1.8)
```

Hmmmm.  That didn't seem to work properly.  Let's handle the "missing values" by removing them (na == "not available" or "not applicable"; rm == "remove"; therefore na.rm == "remove values 
that are not available).

```{r}
summary_monthly_temp <- weather %>% 
  group_by(month) %>% 
  summarize(mean_temp_in_F = mean(temp, na.rm = TRUE), 
            mean_temp_in_C = mean(temp_in_C, na.rm = TRUE))
summary_monthly_temp
```

Calculate a new variable "gain" that is the the difference of dep_delay and arr_delay:

```{r}
flights <- flights %>% 
  mutate(gain = dep_delay - arr_delay)
```



```{r}
gain_summary <- flights %>% 
  summarize(
    min = min(gain, na.rm = TRUE),
    q1 = quantile(gain, 0.25, na.rm = TRUE),
    median = quantile(gain, 0.5, na.rm = TRUE),
    q3 = quantile(gain, 0.75, na.rm = TRUE),
    max = max(gain, na.rm = TRUE),
    mean = mean(gain, na.rm = TRUE),
    sd = sd(gain, na.rm = TRUE),
    missing = sum(is.na(gain))
  )
gain_summary
```


```{r}
ggplot(data = flights, mapping = aes(x = gain)) +
  geom_histogram(color = "white", bins = 20) +
  labs(caption="Histogram of gain variable")
```



```{r}
flights <- flights %>% 
  mutate(
    gain = dep_delay - arr_delay,
    hours = air_time / 60,
    gain_per_hour = gain / hours
  )
```


## (LC4.11) 
Could we create the dep_delay and arr_delay columns by simply subtracting dep_time from sched_dep_time and similarly for arrivals? Take a look at the values in those columns and comment on some challenges that you may encounter.

INSERT YOUR ANSWER HERE
You would get negative numbers correct? It may be more prudent to subtract sched_dep_time from dep_time and same for arrivals.

# Sorting (arrange)

```{r}
freq_dest <- flights %>% 
  group_by(dest) %>% 
  summarize(num_flights = n())
freq_dest
```

```{r}
freq_dest %>% 
  arrange(num_flights)
```

```{r}
freq_dest %>% 
  arrange(desc(num_flights))
```


## (LC 544.3.2)
Which airport from the ```airports``` dataframe has the highest altitude?

```{r}
# INSERT YOUR R CODE HERE
airports %>% 
  arrange(desc(alt))
```


## Joining (inner_join)


```{r}
flights_joined <- flights %>%
 inner_join(airlines, by = "carrier")
``` 





```{r}
flights_with_airport_names <- flights %>%
 inner_join(airports, by = c("dest" = "faa"))
``` 


```{r}
named_dests <- flights %>%
  group_by(dest) %>%
  summarize(num_flights = n()) %>%
  arrange(desc(num_flights)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  rename(airport_name = name)
named_dests
```

```{r}
flights_weather_joined <- flights %>%
 inner_join(weather, by = c("year", "month", "day", "hour", "origin"))
```

```{r}
joined_flights <- flights %>%
 inner_join(airlines, by = "carrier")
```

## (LC 544.3.3)
What's the name of the airport that has the largest mean departure delay from any of the New York airports?
Hint: Copy the code above where we created named_dests and modify it.

```{r}
# INSERT YOUR R CODE HERE
largest_dep_delay <- flights %>%
  group_by(origin) %>%
  summarize(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  arrange(desc(mean_dep_delay)) %>%
  inner_join(airports, by = c("origin" = "faa")) %>%
  rename(airport_name = name)
largest_dep_delay

```


## Selecting

```{r}
flights %>%
 select(carrier, flight)
```

```{r}
flights_no_year <- flights %>%
 select(-year)
```


```{r}
flight_arr_times <- flights %>%
 select(month:day, arr_time:sched_arr_time)
```


```{r}
flights_reorder <- flights %>%
 select(year, month, day, hour, minute, time_hour, everything())
```

```{r}
flights_begin_a <- flights %>%
 select(starts_with("a"))
```

```{r}
flights_delays <- flights %>%
  select(ends_with("delay"))
```


```{r}
flights_time <- flights %>%
 select(contains("time"))
```

## Renaming

```{r}
flights_time_new <- flights %>%
 select(contains("time")) %>%
 rename(departure_time = dep_time,
        arrival_time = arr_time)
```

## Tidying

```{r}
library(fivethirtyeight)
```

```{r}
drinks
```

```{r}
drinks_smaller <- drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia")) %>% 
  select(-total_litres_of_pure_alcohol) %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings)
drinks_smaller
```

## (LC 544.3.4)
Add Canada to the list of countries in our drinks_smaller dataframe.  You might just want to copy and paste the above code block and edit it instead of trying to actually add Canada to the existing drinks_smaller dataframe.



```{r}
# INSERT YOUR R CODE HERE
drinks_smaller <- drinks %>% 
  filter(country %in% c("USA", "China", "Italy", "Saudi Arabia", "Canada")) %>% 
  select(-total_litres_of_pure_alcohol) %>% 
  rename(beer = beer_servings, spirit = spirit_servings, wine = wine_servings)
drinks_smaller
```

```{r}
drinks_smaller_tidy <- drinks_smaller %>% 
  gather(key = type, value = servings, c(beer, spirit, wine))
drinks_smaller_tidy
```

```{r}
ggplot(drinks_smaller_tidy, 
       aes(x = country, y = servings, fill = type)) +
  geom_col(position = "dodge")
```

END OF NOTEBOOK
