---
title: "SI 544 Day 4: Descriptive Statistics and Basic Probability"
author: "Matthew Garvin (mgarvin)"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
weight <- c(115, 115, 117, 120, 123, 126, 129, 132, 135, 139, 142, 146, 150, 154, 159, 164, 180)
height <- c(58, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72) 
```

# Central Tendency

## Mean

```{r}
weight_mean <- mean(weight)
height_mean <- mean(height)

sprintf("Mean Weight: %1.4f, Mean Height: %1.4f", weight_mean, height_mean)
```

## Median

```{r}
weight_median <- median(height)
height_median <- median(weight)

sprintf("Median Weight: %1.4f, Median Height: %1.4f", weight_median, height_median)
```


## Minimum

```{r}
min(weight)
```

## Maximum

```{r}
max(weight)
```

## Quantiles

```{r}
quantile(weight)
```

```{r}
quantile(weight, prob = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0))
```
Equivalently:

```{r}
quantile(weight, prob = seq(0, 1, length = 11))
```

```{r}
?quantile
```

## Mode

```{r}
get_mode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

height_mode <- get_mode(height)
weight_mode <- get_mode(weight)

sprintf("Mode Weight: %s, Height Mode: %s", weight_mode, height_mode)
```

# Dispersion

## Range

```{r}
range(weight)
```

## Interquartile Range

```{r}
IQR(weight)
```

## Standard Deviation

```{r}
weight_sd <- sd(weight)
height_sd <- sd(height)

sprintf("Weight SD: %1.4f, Height SD: %1.4f", weight_sd, height_sd)
```

## z-scores

One way:

```{r}
x <- 70

zWeight <- (x - mean(weight) ) / sd(weight)
zHeight <- (x - mean(height) ) / sd(height)
sprintf("Weight Z: %1.2f. Height Z: %1.2f", zWeight, zHeight)

```

Another way:

```{r}

zWeight <- scale(x, center = mean(weight), scale = sd(weight))
zHeight <- scale(x, center = mean(height), scale = sd(height))
sprintf("Weight Z: %1.2f. Height Z: %1.2f", zWeight, zHeight)
```

What's the probability of someone being less than 74" tall?

```{r}

pnorm(74, mean(height), sd(height))    #using x, mu, and sigma

```

So what's the probability of someone being greater than 74" tall?

```{r}
1 - pnorm(74, mean(height), sd(height)) 
```

# Putting our knowledge together

Let's load some additional data:

```{r}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")
```

The description of the data is available from https://www.openintro.org/data/index.php?data=bdims

## Learning Check 544.4.1

Use pipes to create two new dataframes (or tibbles): 
one called fdims containing only those measurements for "females" and 
one called mdims containing only those measurements for "males". 

Note that gender is no longer considered a binary variable, but the dataset was collected without
such information.

Hint: see "filter" from last lecture.

```{r}
library(tidyverse)
```

```{r}

```


```{r}
# INSERT YOUR R CODE HERE
fdims <- bdims %>%
  filter(sex == 0)

mdims <- bdims %>%
  filter(sex == 1)
```

## Learning Check 544.4.2

Create a histogram for the distribution of heights for females.  Make a good choice for binwidth or bins.

```{r}
# INSERT YOUR R CODE HERE
fdims_histo <- ggplot(fdims, mapping = aes(x=hgt)) +
  geom_histogram(bins=15)
fdims_histo
```

## Learning Check 544.4.3

Calculate the mean and standard deviation for the heights of females.

```{r}
# INSERT YOUR R CODE HERE

sprintf("Mean Height: %1.4f, Standard Deviation: %1.4f", mean(fdims$hgt), sd(fdims$hgt))


```


## Learning Check 544.4.4

Comment on the relationship between the histogram you generated and the computed values of the mean and 
standard deviation.

Mean is average heights of all female. It looks to be in the middle of the two peaks. Most of the females lie within +/- 19.5 cm, which is 3 times the standard deviation.



Let's take a look at a slightly richer histogram.  This one overlays a z-distribution that has a mean and standard
deviation equal to that of the heights from fdims:

```{r}
ggplot(fdims, mapping=aes(x=hgt)) + 
  geom_histogram(aes(y = ..density..), bins = 15, fill="blue") + 
  stat_function(fun=dnorm, args=with(fdims, c(mean=mean(hgt), sd = sd(hgt))))
```
```{r}

pnorm(180, mean(mdims$hgt), sd(mdims$hgt))    #using x, mu, and sigma

```

## Learning Check 544.4.5

What's the probability that any given male person is at least 180cm tall?

Probability that any given male person is at least 180 cm tall is 0.6231864, or 62.3%

