---
title: "si544-day5-basic-regression"
author: "Matthew Garvin"
uniqename: "mgarvin"
date: "10/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
library(gapminder)
```

## Load & Look


```{r}
evals_ch6 <- evals %>%
  select(ID, score, bty_avg, age)
```

```glimpse``` can be helpful:

```{r}
glimpse(evals_ch6)
```

Alternatively, you can draw a random sample of some values and look at what you get:

```{r}
evals_ch6 %>% sample_n(5)
```

Now that you understand descriptive statistics and visualizations, let's take a shortcut:


```{r}
evals_ch6 %>%
  select(score, bty_avg) %>%
  skim()

```

```{r}
evals_ch6 %>%
  get_correlation(formula = score ~ bty_avg)
```

Equivalently:

```{r}
evals_ch6 %>%
  summarize(correlation = cor(score, bty_avg))
```

Visualization:

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_jitter() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Scatterplot of relationship of teaching and beauty scores")
```

```{r}
ggplot(evals_ch6, aes(x = bty_avg, y = score)) +
  geom_point() +
  labs(x = "Beauty Score", y = "Teaching Score",
       title = "Relationship between teaching and beauty scores") +  
  geom_smooth(method = "lm", se = FALSE)
```

### Learning Check 544.5.1

Repeat this for our handwriting data, looking at the relationship between dominant and non_dominant.  The handwriting dataset is available via
https://raw.githubusercontent.com/umsi-data-science/data/main/handwriting.csv (remember that you can give read_csv a URL).

First, load the data into a variable called ```handwriting```.

```{r}
# INSERT YOUR R CODE HERE
handwriting <- read.csv("https://raw.githubusercontent.com/umsi-data-science/data/main/handwriting.csv")
```

1. Use glimpse() to get a "feel" for the handwriting data.

```{r}
# INSERT YOUR R CODE HERE
glimpse(handwriting)
```

2. Use skim() to generate summary statistics for dominant and non-dominant times.

```{r}
# INSERT YOUR R CODE HERE
handwriting %>%
  select(dominant, nondominant) %>%
  skim()
```

3. Calculate the correlation coefficient for dominant and non-dominant times.  What does that value tell you?

```{r}
# INSERT YOUR R CODE HERE
handwriting %>% 
  summarise(corr_coef = cor(dominant, nondominant))
```

The correlation coefficient is .705

4. Create a scatterplot with non-dominant on the y (vertical) axis and dominant on the x (horizontal) axis.  Use R to draw a line through the data to show the relationship between dominant and non-dominant times.

```{r}
# INSERT YOUR R CODE HERE
ggplot(handwriting, aes(dominant, nondominant)) +
  geom_point() +
    labs(x = "Dominant", y = "Nondominant",
       title = "Relationship between dominant and nondominant") +  
  geom_smooth(method = "lm", se = FALSE)
```


## Simple Linear Regression

```{r}
# Fit regression model:
score_model <- lm(score ~ bty_avg, data = evals_ch6)
```


```{r}
# Get regression table:
get_regression_table(score_model)
```

```{r}
get_regression_summaries(score_model)
```

### Best fit

```{r}
regression_points <- get_regression_points(score_model)
regression_points
```

Let's calculate the sum of squared residuals using a pipeline:

```{r}
regression_points %>%
  mutate(squared_residuals = residual^2) %>%
  summarize(sum_of_squared_residuals = sum(squared_residuals))

```


```{r}
ggplot(regression_points, mapping=aes(x=residual)) + geom_histogram()
```


### Learning Check 544.5.2

Repeat the above steps with the handwriting data from discussion section:

1. Create a linear model (using the lm() function) to model non-dominant time as the outcome variable and dominant time as the explanatory one.

```{r}
# INSERT YOUR R CODE HERE
handwriting_model <- lm(nondominant ~ dominant, data = handwriting)
```

2. Use get_regression_table() to get the coefficients for the regression model.

```{r}
# INSERT YOUR R CODE HERE
get_regression_table(handwriting_model)
```


3. What would the expected value of non-dominant time be for a name that took 6 seconds to write with one's dominant hand?
3.129 + 1.853 * 6 

We can predict 14.247 seconds to write your name with your nondominant hand.

4. Use get_regression_summaries() to get information about how "good" the model is.  Comment on how much variance is explained by the model.  What's the average deviation from the predicted value?

```{r}
# INSERT YOUR CODE HERE
get_regression_summaries(handwriting_model)
```

.49 is the variance, and 1.67 is the average deviation.

5. Identify the two most extreme residual values.  What do those values mean?  Can you infer anything about the individuals with those values?

```{r}
# INSERT YOUR R CODE HERE
regression_points <- get_regression_points(handwriting_model) %>% 
  summarise(max_residual = max(residual, na.rm = TRUE), min_residual = min(residual, na.rm = TRUE))
regression_points

get_regression_points(handwriting_model) %>% arrange(desc(residual))
get_regression_points(handwriting_model) %>% arrange(residual)
```

max residual = 12.694, min residual = -10.662
It would seem that we can infer that the individual with the max residual is ambidextrous.

