---
title: "si544-day7-multiple-regression"
author: "Matthew Garvin (mgarvin)"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As usual, load all the libraries that we'll need:

```{r}
library(tidyverse)
library(moderndive)
library(skimr)
```
Let's load the Szeged weather data:

```{r}
weather <- read_csv("https://github.com/umsi-data-science/data/raw/main/weatherHistory.csv")
```

## Learning Check 544.8.1: 

Create a dataframe called "weather_day7" based on a random sample of 500 rows from weather that consists of only the following columns:
temp (renamed from "Temperature (C)" ), apparent_temp (renamed from "Apparent Temperature ( C)", humidity (from Humidity), and wind_speed (from "Wind Speed (km/h)"


```{r}
# INSERT YOUR R CODE HERE
set.seed(42)
weather_day7 <- weather %>%
  select(temp = `Temperature (C)`, apparent_temp = `Apparent Temperature (C)`, humidity = Humidity, wind_speed = `Wind Speed (km/h)`) %>%
  sample_n(500)

```

Let's take a look at the relationship between apparent_temp and temp, as well as apparent_temp
and humidity:

```{r}
ggplot(weather_day7, mapping=aes(x=temp, y=apparent_temp)) + geom_point()
ggplot(weather_day7, mapping=aes(x=humidity, y=apparent_temp)) + geom_point()
```

## Learning Check 544.8.2:

Create a regression that models "Apparent Temperature" in terms of "Temperature".  How good is your model?

```{r}
# INSERT YOUR R CODE HERE
r_model <- lm(apparent_temp ~ temp, data = weather_day7)
# Get regression table:
get_regression_table(r_model)
get_regression_summaries(r_model)
```

INSERT YOUR INTERPRETATION HERE.


Let's look at a model of apparent_temperature that uses temp and humidity as explanatory
variables:

```{r}
weather_model <- lm(apparent_temp ~ temp*humidity
                   ,data=weather_day7)
get_regression_table(weather_model)
get_regression_summaries(weather_model)
```


## Learning Check 544.8.3:

Create a linear regression model for apparent_temp that includes temperature, humidity, and wind
speed as explanatory variables.  Calculate the apparent temperature when the temperature is
20 degrees Celsius, the humidity 65%, and the wind speed 20 km/h.

```{r}
# INSERT YOUR R CODE HERE
weather_model2 <- lm(apparent_temp ~ temp*humidity*wind_speed
                   ,data=weather_day7)
get_regression_table(weather_model2)
get_regression_summaries(weather_model2)
```

```{r}
-1.505 + 1.029*20 + 1.580*.65 +20*20*(-.005) + .65*20*(-0.359) + 20*.65*20*0.024
```


INSERT YOUR ANSWER HERE
19.675

